---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rstanemax

[![Travis-CI Build Status](https://travis-ci.org/yoshidk6/rstanemax.svg?branch=master)](https://travis-ci.org/yoshidk6/rstanemax)
[![CRAN status](https://www.r-pkg.org/badges/version/rstanemax)](https://cran.r-project.org/package=rstanemax)
[![downloads](https://cranlogs.r-pkg.org/badges/grand-total/rstanemax)](https://CRAN.R-project.org/package=rstanemax)


This small package performs simple sigmoidal Emax model fit using Stan, without the need of (1) writing Stan model code and (2) setting up an environment to compile Stan model, inspired by __rstanarm__ package.

__rstanarm__ package ([link](https://mc-stan.org/users/interfaces/rstanarm)) is a very flexible, general purpose tool to perform various Bayesian modeling with formula notations, such as generalized mixed effect models or joint models.
One small gap it has is in nonlinear model fitting, where it only accepts nonlinear functions defined in stats package with `SS` prefixes ([link](http://mc-stan.org/rstanarm/articles/glmer.html#relationship-to-nlmer)). 
Unfortunately the (sigmoidal) Emax model, one of the most commonly used nonlinear functions in the field of pharmacometrics, is not among the available functions.
The __rstanarm__ package also seems to be assuming that we fit nonlinear mixed effect models, but not simple nonlinear models without mixed effects. 
__brms__ is another great general purpose package that allows formula notation, but it requires users to compile generated Stan models. The need for setting up an environment for compiling Stan code can be daunting for beginners, especially with Windows.

I hope this __rstanemax__ package will fill the niche gap, allow for easier implementation of Emax model in Bayesian framework, and enable routine uses in the pharmacokinetic/pharmacodynamic field.

This package was build using __rstantools__ ([link](https://mc-stan.org/rstantools/)) following a very helpful step-by-step guide ([link](https://mc-stan.org/rstantools/articles/minimal-rstan-package.html)) on creating a package that depends on RStan. 


## Installation

### From CRAN

You can install the released version of rstanemax from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("rstanemax")
```

### From source

You can alternatively install the package from source.  
Before doing so, you first have to install RStan and C++ Toolchain.  
[RStan Getting Started](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started)  
Also, you have to follow the instruction below if you are using Windows PC.  
[Installing RStan from source on Windows](https://github.com/stan-dev/rstan/wiki/Installing-RStan-from-source-on-Windows)  

After this step you should be able to install the package from GitHub using __devtools__.

``` r
install.packages("devtools")
library(devtools)
devtools::install_github("yoshidk6/rstanemax")
```

## Example

[This GitHub pages](https://yoshidk6.github.io/rstanemax) contains function references and [vignette](https://yoshidk6.github.io/rstanemax/articles/emaxmodel.html).

```{r}
# Load rstanemax
library(rstanemax)

# Run model with a sample dataset
set.seed(12345)

data(exposure.response.sample)

fit.emax <- stan_emax(response ~ exposure, data = exposure.response.sample,
                      # the next line is only to make the output short
                      chains = 1, iter = 500, seed = 12345)
```

```{r}
fit.emax
```

```{r}
plot(fit.emax)
```

