---
title: "Simple Emax model fit with Stan"
author: "Kenta Yoshida"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple Emax model fit with Stan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE}
library(rstanemax)
library(dplyr)
library(ggplot2)
set.seed(12345)
```

This vignette provide an overview of the workflow of Emax model analysis using this package.


# Typical workflow

## Model run with `stan_emax` function

`stan_emax()` is the main function of this package to perform Emax model analysis on the data.
This function requires minimum two input arguments - `formula` and `data`.
In the `formula` argument, you will specify which columns of `data` will be used as exposure and response data, in a format similar to `stats::lm()` function, e.g. `response ~ exposure`.

```{r, results="hide"}
data(exposure.response.sample)

fit.emax <- stan_emax(response ~ exposure, data = exposure.response.sample)
```

```{r}
fit.emax
```

`plot()` function shows the estimated Emax model curve with 95% credible intervals of parameters.

```{r, fig.show='hold'}
plot(fit.emax)
```

Raw output from `rstan` is stored in the output variable, and you can access it with `$stanfit`.

```{r}
class(fit.emax$stanfit)
```

## Prediction of response with new exposure data

`posterior_predict()` function allows users to predict the response using new exposure data.
If `newdata` is not provided, the function returns the prediction on the exposures in original data.
The default output is a matrix of posterior predictions, but you can also specify "dataframe" or "tibble" that contain posterior predictions in a long format.
See help of `rstanemax::posterior_predict()` for the description of two predictions, `respHat` and `response`.

```{r}
response.pred <- posterior_predict(fit.emax, newdata = c(0, 100, 1000), returnType = "tibble")

response.pred %>% select(mcmcid, exposure, respHat, response)
```

You can also get quantiles of predictions with `posterior_predict_quantile()` function.
Currently this function only export 95% credible intervals of mean response and 95% prediction intervals of individual responses.

```{r}
resp.pred.quantile <- posterior_predict_quantile(fit.emax, newdata = seq(0, 1000, by = 250))
resp.pred.quantile
```

This is particularly useful when you want to plot the estimated Emax curve.


```{r}
ggplot(resp.pred.quantile, aes(exposure, ci_med)) +
  geom_line() + 
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high), alpha = .5) +
  geom_ribbon(aes(ymin=pi_low, ymax=pi_high), alpha = .2) +
  labs(y = "response")
```



# Fix parameter values in Emax model

You can fix parameter values in Emax model  for E0 and/or gamma (Hill coefficient).
See help of `stan_emax()` for the details.
The default is to fix gamma at 1 and to estimate E0 from data.

Below is the example of estimating gamma from data.

```{r, results="hide"}
data(exposure.response.sample)

fit.emax.sigmoidal <- stan_emax(response ~ exposure, data = exposure.response.sample, gamma.fix = NULL)
```

```{r}
fit.emax.sigmoidal
```

You can compare the difference of posterior predictions between two models (in this case they are very close to each other):

```{r fig.width = 6, fig.height = 4}

exposure_pred <- seq(min(exposure.response.sample$exposure),
                     max(exposure.response.sample$exposure),
                     length.out = 100)

pred1 <- 
  posterior_predict_quantile(fit.emax, exposure_pred) %>% 
  mutate(model = "Emax")
pred2 <- 
  posterior_predict_quantile(fit.emax.sigmoidal, exposure_pred) %>% 
  mutate(model = "Sigmoidal Emax")

pred <- bind_rows(pred1, pred2)


ggplot(pred, aes(exposure, ci_med, color = model, fill = model)) +
  geom_line() + 
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high), alpha = .3) +
  geom_ribbon(aes(ymin=pi_low, ymax=pi_high), alpha = .1, color = NA) +
  geom_point(data=exposure.response.sample, aes(exposure, response), 
             color = "black", fill = NA, size=2) +
  labs(y = "response")



 
```





