---
title: "Simple Emax model fit with Stan"
author: "Kenta Yoshida"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple Emax model fit with Stan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE}
library(rstanemax)
library(dplyr)
library(ggplot2)
set.seed(12345)
```

This vignette provide an overview of the workflow of Emax model analysis using this package.


# Typical workflow

## Model run with `stan_emax` function

`stan_emax()` is the main function of this package to perform Emax model analysis on the data.
This function requires minimum two input arguments - `formula` and `data`.
In the `formula` argument, you will specify which columns of `data` will be used as exposure and response data, in a format similar to `stats::lm()` function, e.g. `response ~ exposure`.

```{r, results="hide"}
data(exposure.response.sample)

fit.emax <- stan_emax(response ~ exposure, data = exposure.response.sample,
                      # the next line is only to make the example go fast enough
                      chains = 2, iter = 1000, seed = 12345)
```

```{r}
fit.emax
```

`plot()` function shows the estimated Emax model curve with 95% credible intervals of parameters.

```{r plot_example, fig.show='hold'}
plot(fit.emax)
```

Output of `plot()` function (for `stanemax` object) is a `ggplot` object, so you can apply additional settings as you would do in `ggplot2`.  
Here is an example of using log scale for x axis (note that exposure == 0 is hanging at the very left, making the curve a bit weird).

```{r plot_example_log, fig.show='hold'}
plot(fit.emax) + scale_x_log10() + expand_limits(x = 1)
```


Raw output from `rstan` is stored in the output variable, and you can access it with `extract_stanfit()` function.

```{r}
class(extract_stanfit(fit.emax))
```

## Prediction of response with new exposure data

`posterior_predict()` function allows users to predict the response using new exposure data.
If `newdata` is not provided, the function returns the prediction on the exposures in original data.
The default output is a matrix of posterior predictions, but you can also specify "dataframe" or "tibble" that contain posterior predictions in a long format.
See help of `rstanemax::posterior_predict()` for the description of two predictions, `respHat` and `response`.

```{r}
response.pred <- posterior_predict(fit.emax, newdata = c(0, 100, 1000), returnType = "tibble")

response.pred %>% select(mcmcid, exposure, respHat, response)
```

You can also get quantiles of predictions with `posterior_predict_quantile()` function.

```{r}
resp.pred.quantile <- posterior_predict_quantile(fit.emax, newdata = seq(0, 5000, by = 100))
resp.pred.quantile
```

Input data can be obtained in a same format with `extract_obs_mod_frame()` function.

```{r}
obs.formatted <- extract_obs_mod_frame(fit.emax)
```

These are particularly useful when you want to plot the estimated Emax curve.

```{r plot_with_pp, fig.show='hold'}
ggplot(resp.pred.quantile, aes(exposure, ci_med)) +
  geom_line() + 
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high), alpha = .5) +
  geom_ribbon(aes(ymin=pi_low, ymax=pi_high), alpha = .2) +
  geom_point(data = obs.formatted,
             aes(y = response)) +
  labs(y = "response")
```



# Fix parameter values in Emax model

You can fix parameter values in Emax model for Emax, E0 and/or gamma (Hill coefficient).
See help of `stan_emax()` for the details.
The default is to fix gamma at 1 and to estimate Emax and E0 from data.

Below is the example of estimating gamma from data.

```{r, results="hide"}
data(exposure.response.sample)

fit.emax.sigmoidal <- stan_emax(response ~ exposure, data = exposure.response.sample, 
                                gamma.fix = NULL,
                                # the next line is only to make the example go fast enough
                                chains = 2, iter = 1000, seed = 12345)
```

```{r}
fit.emax.sigmoidal
```

You can compare the difference of posterior predictions between two models (in this case they are very close to each other):

```{r plot_with_gamma_fix, fig.width = 6, fig.height = 4, fig.show='hold'}

exposure_pred <- seq(min(exposure.response.sample$exposure),
                     max(exposure.response.sample$exposure),
                     length.out = 100)

pred1 <- 
  posterior_predict_quantile(fit.emax, exposure_pred) %>% 
  mutate(model = "Emax")
pred2 <- 
  posterior_predict_quantile(fit.emax.sigmoidal, exposure_pred) %>% 
  mutate(model = "Sigmoidal Emax")

pred <- bind_rows(pred1, pred2)


ggplot(pred, aes(exposure, ci_med, color = model, fill = model)) +
  geom_line() + 
  geom_ribbon(aes(ymin=ci_low, ymax=ci_high), alpha = .3) +
  geom_ribbon(aes(ymin=pi_low, ymax=pi_high), alpha = .1, color = NA) +
  geom_point(data=exposure.response.sample, aes(exposure, response), 
             color = "black", fill = NA, size=2) +
  labs(y = "response")



 
```



# Set covariates

You can specify categorical covariates for Emax, EC50, and E0.
See help of `stan_emax()` for the details.


```{r, results="hide"}
data(exposure.response.sample.test)

test.data <-
  mutate(exposure.response.sample.test,
         SEX = ifelse(cov2 == 0, "MALE", "FEMALE"))

fit.cov <- stan_emax(formula = resp ~ conc, data = test.data,
                     param.cov = list(emax = "SEX"),
                     # the next line is only to make the example go fast enough
                     chains = 2, iter = 1000, seed = 12345)
```

```{r plot_with_cov, fig.width = 6, fig.height = 4, fig.show='hold'}
fit.cov
plot(fit.cov)
```


You can extract MCMC samples from raw stanfit and evaluate differences between groups.

```{r compare_emax, fig.show='hold'}
emax.posterior <- 
  fit.cov %>% 
  extract_stanfit() %>% 
  rstan::extract(pars = "emax") %>% 
  .$emax

# delta = emax[FEMALE] - emax[MALE]
delta.emax.posterior <- emax.posterior[,1] - emax.posterior[,2]

ggplot2::qplot(delta.emax.posterior, bins = 30) +
  ggplot2::labs(x = "emax[FEMALE] - emax[MALE]")

# Credible interval of delta
quantile(delta.emax.posterior, probs = c(0.025, 0.05, 0.5, 0.95, 0.975))

# Posterior probability of emax[FEMALE] < emax[MALE]
sum(delta.emax.posterior < 0) / length(delta.emax.posterior)

```


